---
title: "Motion cloud masks and targets"
author: "Egor Ananyev"
date: "July 3, 2018"
output:
    html_document:
        toc: true
---

<style type="text/css">
body, td{
  font-family: Cambria;
}
code.r{
  font-family: Inconsolata;
}
</style>

# Preparing the data

## Loading the packages and data directories
```{r set-options, message=F}
options(width=100)
out=F # do we output the plots?
render=T # do we render the plots?
qc=T # do we need quality control (stair plots)?
an=T # do we run the analyses?
# The list of subjects, the order of conditions, and the thresholds are derived from Subjects.xlsx
library(xlsx)
library(plyr)
library(reshape)
library(matrixStats)
#library(splines)
db <- '/home/egor/Dropbox/' # on Linux
db <- '/Users/Egor/Dropbox/' # Windows
# db <- '~/Dropbox/' # on Mac
if(out || render){
    library(ggplot2)
    library(gridExtra)
}
if(an){
    source(paste(db, 'Prog/R/myFunctions/pvalfn.R', sep=''))
    library(lme4)
    library(lmerTest)
    library(BayesFactor)
}
```

## Plot variables
```{r, message=FALSE, warning=FALSE}
if(out || render){
    # theme for plotting:
    alpha <- .7
    w <- .56 # proportion width in group plots
    xLab <- 'Target Peak Time (s)'
    yLab <- 'Log Contrast Threshold'
    # colLab <- expression(paste('\nTarget\nVelocity (', degree, '/s)', sep=''))
    colLab <- expression(paste('Target Eccentricity (', degree, ')', sep=''))
    # colLabType <- 'Mask Type'
    yLim <- c(-1.7,-0.5)
    dodge <- position_dodge(width=0.1)
}
```

## Plot functions
```{r}
if(out || render){
    themefy <- function(p) {
        p <- p + theme_bw() + 
             theme(panel.grid.minor.x=element_blank(), panel.grid.minor.y=element_blank(),
                axis.text=element_text(size=8), axis.title=element_text(size=9),
                legend.text=element_text(size=8), legend.title=element_text(size=9),
                legend.key = element_blank(), legend.margin=margin(t=-.04, unit='in'),
                legend.background = element_rect(fill='transparent'),
                plot.title=element_text(face='bold'))
    }
    sumFn <- function(ss, subjStr='subj', xStr='targTpeak', grpStr='targEcc'){
        sumSubj <- ddply(ss, c(subjStr, xStr, grpStr), summarise,
                         mnS=mean(threshMean), se=sd(threshMean)/sqrt(length(threshMean))) 
        # total mean across conditions per subj
        sumSubjMn <- ddply(ss, c(subjStr), summarise, mnStot=mean(threshMean)) 
        sumSubj <- merge(sumSubj, sumSubjMn)
        sumSubj$normS <- - sumSubj$mnS / sumSubj$mnStot # normalized subject mean
        sumSubj$seNorm <- NA
        # sumSubj$mnS[is.na(sumSubj$mnS)] <- 0
        sumGrp <- ddply(sumSubj, c(xStr, grpStr), summarise,
                      mn=mean(mnS), se=sd(mnS)/sqrt(length(mnS)),
                      norm=mean(normS), seNorm=sd(normS)/sqrt(length(normS)))
        sumGrp$subj <- 'average'
        sumSubj <- rename(sumSubj, c(mnS='mn',normS='norm'))
        sumComb <- rbind(sumGrp, subset(sumSubj, select=-mnStot))
        sumComb$se[is.na(sumComb$se)] <- 0
        sumComb
    }
    plotAve <- function(pss, subjStr='subj', xStr='targTpeak', grpStr='targEcc', 
                        xlab=xLab, ylab=yLab, collab=colLab, yStr='mn', seStr='se'){
        pss$yMin <- pss[,yStr] - pss[,seStr]
        pss$yMax <- pss[,yStr] + pss[,seStr]
        pss[,grpStr] <- factor(pss[,grpStr])
        p <- ggplot(pss, aes_string(x=xStr, y=yStr, colour=grpStr, group=grpStr,
                                ymin='yMin', ymax='yMax')) +
            geom_point(position=dodge, size=1, alpha=alpha) + 
            geom_line(position=dodge, alpha=alpha) +
            scale_x_continuous(breaks=c(.5,1,1.5), labels=c('0.5','1','1.5')) +
            geom_linerange(position=dodge, show.legend=F, alpha=alpha) +
            labs(x=xlab, y=ylab, colour=collab) + ylim(yLim) +
            guides(colour=guide_legend(keyheight=.3, default.unit='inch'))
        p <- themefy(p)
    }
    plotIndiv <- function(pss, subjStr='subj', xStr='targTpeak', grpStr='targEcc', 
                        xlab=xLab, ylab=yLab, collab=colLab, yStr='mn', seStr='se'){
        pss$yMin <- pss[,yStr] - pss[,seStr]
        pss$yMax <- pss[,yStr] + pss[,seStr]
        pss[,grpStr] <- factor(pss[,grpStr])
        p <- ggplot(pss, aes_string(x=xStr, y=yStr, colour=grpStr, group=grpStr,
                                ymin='yMin', ymax='yMax')) + 
            facet_wrap( ~ subj, ncol=4) +
            geom_point(position=dodge, size=1, alpha=alpha) + 
            geom_line(position=dodge, alpha=alpha) +
            geom_linerange(position=dodge, show.legend=F, alpha=alpha) +
            scale_x_continuous(breaks=c(.5,1,1.5), labels=c('0.5','1','1.5')) +
            labs(x=xlab, y=ylab, colour=collab) + ylim(-2,0) + 
            guides(colour=guide_legend(keyheight=.3, default.unit='inch'))
        p <- themefy(p)
    }
}
```

## Loading the data
```{r}
allDataDir <- paste(db,'Projects/mc/data_mc3',sep='')
dataDirs <- dir(allDataDir)
dataDirs <- dataDirs[grep('mc3',dataDirs)]
colsOfInt <- c('participant', 'dom', 'session', 'mcBv', 'targTpeak', 'targXoff2', 
               'stairStart', 'meanRev6')
df <- data.frame()
dfRevs <- data.frame()
dfIntn <- data.frame()
for(curDir in dataDirs){
    print(curDir)
    curDf <- read.csv(paste(allDataDir,'/',curDir,'/',curDir,'.csv',sep=''))
    curDf <- curDf[,colsOfInt]
    if(qc){
        subjStairs <- dir(paste(allDataDir,'/',curDir,'/',sep=''))
        subjStairs <- subjStairs[grep('.tsv',subjStairs)]
        for(curStairFN in subjStairs){
            curStair <- paste(allDataDir,'/',curDir,'/',curStairFN,sep='')
            curRevs <- read.table(curStair, skip=1, nrows=1)
            curIntn <- read.table(curStair, skip=4, nrows=2)
            curInfo <- readLines(curStair)
            curInfoCols <- data.frame(subj=curDf$participant[1], dom=curDf$dom[1], 
                                      sess=curDf$session[1],
                                      stairStart=curInfo[which(curInfo==" u'startContr':")+1], 
                                      mcBv=curInfo[which(curInfo==" u'mcBv':")+1], 
                                      targTpeak=curInfo[which(curInfo==" u'targTpeak':")+1], 
                                      targEcc=curInfo[which(curInfo==" u'targXoff2':")+1])
            curDfRevs <- cbind(curInfoCols, curRevs[,2:9])
            nTrials <- ncol(curIntn)-1
            curDfIntn <- curInfoCols[rep(seq_len(nrow(curInfoCols)), each=nTrials),]
            curDfIntn$trial <- seq(1,nTrials)
            curDfIntn$intn <- as.numeric(curIntn[1,2:(nTrials+1)])
            curDfIntn$resp <- as.numeric(curIntn[2,2:(nTrials+1)])
            rownames(curDfIntn) <- NULL
            dfRevs <- rbind(dfRevs, curDfRevs)
            dfIntn <- rbind(dfIntn, curDfIntn)
        }
    }
    df <- rbind(df, curDf)
}
if(qc){
    dfIntn$maskV <- round(as.numeric(levels(dfIntn$mcBv))[dfIntn$mcBv] * 60 / 35, 1)
    dfIntn$maskV[dfIntn$maskV<0.05] <- 0
    head(dfIntn)
}
ds <- rename(df, c(participant='subj', session='sess', meanRev6='thresh',
                   mcBv='maskV', targXoff2='targEcc'))
ds$targEcc <- round(ds$targEcc / 35,1)
ds$maskType <- ''
ds$maskType[ds$maskV==0.01] <- 'stationary'
ds$maskType[ds$maskV==9.6] <- 'fast'
ds$maskV <- round(ds$maskV * 60 / 35, 1)
ds$maskV[ds$maskV<0.05] <- 0
head(ds)
```

## Transformed data sets

Collapsing across low and high stair starts to derive the averages for each condition.
```{r}
thresh <- ddply(ds, .(subj,dom,sess,maskV,maskType,targTpeak,targEcc), summarise, 
                threshMean = mean(thresh))
head(thresh)
```

# QC
For quality control, it's best not to include the description of the conditions in the plots, as the convergence of the staircases should be the main focus.
```{r, fig.height=10, fig.width=10}
if(qc){
    plotQc <- function(pss){
        plot(ggplot(pss, aes(x=trial, y=intn, colour=targEcc, linetype=stairStart)) +
            facet_wrap(subj ~ targTpeak, ncol=6) + geom_point() + geom_line() + ylim(-2,0))
    }
    a <- dlply(dfIntn, .(maskV), plotQc) 
}
```

# Individual plots
```{r, fig.height=6, fig.width=7}
if(render){
    plotIndivFn <- function(pss){
        ss <- sumFn(pss)
        plotSs <- ss[ss$subj!='average',]
        plot(plotIndiv(plotSs) + ggtitle(paste('maskV: ', unique(pss$maskType))))
    }
    a <- dlply(thresh, .(maskType), plotIndivFn)
}
```

# Group plots
```{r, fig.height=7.2, fig.width=7}
if(render || out){
    groupPlot <- function(pss, maskVstr){
        ss <- sumFn(pss[(pss$maskType==maskVstr),])
        themefy(plotAve(ss[ss$subj=='average',]))
    }
    p <- grid.arrange(groupPlot(thresh,'stationary') + ggtitle('stationary mask'), 
                      groupPlot(thresh,'fast') + ggtitle('fast mask'),
                      ncol=2, nrow=1) #widths=c((1-w)/2,w/2))
    # if(render){plot(p)}
    if(out){
        jpeg('mc3.jpg', width=7, height=6.3, units='in', res=600)
        plot(p)
        def.off()
        }
}
```

# Analyses

## Scaling function
```{r}
cent <- function(v){
    v <- apply(v,2,function(x){
        x <- x - mean(unique(x),na.rm=T)
        x <- x / max(x)
    })
}
```

### Centered data set
```{r}
dsc <- ds
centCols <- c('dom','targTpeak','stairStart','targEcc','maskV','targV','sess')
dsc[,centCols] <- cent(dsc[,centCols])
# dsc$maskV_b <- (dsc$maskV + 1) / 2
# dsc$targV_b <- (dsc$targV + 1) / 2
dsc$targEcc_b <- round((dsc$targEcc + 1) / 2,0)
dsc$targTpeak_b <- (dsc$targTpeak + 1) / 2
dsc$targTpeak_c <-  dsc$targTpeak_b * 2 - 1
dsc$targEcc_c <-  dsc$targEcc_b * 2 - 1
dsc$maskType <- factor(dsc$maskType, c('stationary','slow','fast'))
dsc <- dsc[dsc$targV==-1,]
head(dsc)
```

## GLM
```{r}
pvalfn(lmer(thresh ~ dom + stairStart + sess + maskType * targTpeak_b * targEcc_b +
                 (1|subj), data=dsc))
pvalfn(lmer(thresh ~ dom + stairStart + sess + maskType * targTpeak_c * targEcc_c +
                 (1|subj), data=dsc))
```

## BF

```{r}
ss <- dsc
ss$slow <- 0
ss$slow[ss$maskType=='slow'] <- 1
ss$fast <- 0
ss$fast[ss$maskType=='fast'] <- 1
ss$subj <- factor(ss$subj)
```

```{r}
## on stationary
ss_stat <- ss[ss$maskType=='stationary',]
1/generalTestBF(thresh ~ dom + sess + stairStart + targTpeak_b + targEcc_b + subj, 
                  data=ss_stat, whichRandom='subj', neverExclude='subj', 
                  whichModels='top')
mMain <- lmBF(thresh ~ dom + sess + stairStart + targTpeak_b + targEcc_b + subj, 
              data=ss_stat, whichRandom='subj')
mInt <- lmBF(thresh ~ dom + sess + stairStart + targTpeak_b + targEcc_b + subj +
                 targTpeak_b:targEcc_b, 
              data=ss_stat, whichRandom='subj')
mInt/mMain
```

```{r}
## on slow & stationary
ss_nonfast <- ss[ss$maskType!='fast',]
mMain <- lmBF(thresh ~ dom + sess + stairStart + targTpeak_b + targEcc_b + subj +
                  targTpeak_b:targEcc_b, data=ss_nonfast, whichRandom='subj')
mInt <- lmBF(thresh ~ dom + sess + stairStart + targTpeak_b + targEcc_b + subj +
                 targTpeak_b:targEcc_b + slow, 
              data=ss_nonfast, whichRandom='subj')
mInt/mMain
mInt1 <- lmBF(thresh ~ dom + sess + stairStart + targTpeak_b + targEcc_b + subj +
                 targTpeak_b:targEcc_b + slow + slow:targTpeak_b, 
              data=ss_nonfast, whichRandom='subj')
mInt1/mInt
mInt2 <- lmBF(thresh ~ dom + sess + stairStart + targTpeak_b + targEcc_b + subj +
                 targTpeak_b:targEcc_b + slow + slow:targEcc_b, 
              data=ss_nonfast, whichRandom='subj')
mInt2/mInt
```

```{r}
## on slow & stationary
ss_nonslow <- ss[ss$maskType!='slow',]
mMain <- lmBF(thresh ~ dom + sess + stairStart + targTpeak_b + targEcc_b + subj +
                  targTpeak_b:targEcc_b, data=ss_nonslow, whichRandom='subj')
mInt <- lmBF(thresh ~ dom + sess + stairStart + targTpeak_b + targEcc_b + subj +
                 targTpeak_b:targEcc_b + fast, 
              data=ss_nonslow, whichRandom='subj')
mInt/mMain
mInt1 <- lmBF(thresh ~ dom + sess + stairStart + targTpeak_b + targEcc_b + subj +
                 targTpeak_b:targEcc_b + fast + fast:targTpeak_b, 
              data=ss_nonslow, whichRandom='subj')
mInt1/mInt
mInt2 <- lmBF(thresh ~ dom + sess + stairStart + targTpeak_b + targEcc_b + subj +
                 targTpeak_b:targEcc_b + fast + fast:targEcc_b, 
              data=ss_nonslow, whichRandom='subj')
mInt2/mInt
```
